---
title: "Analysis"
author: "Christina Maimone"
date: "4/23/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

This file contains the public data analysis for "Characterizing the US Research Computing and Data (RCD) Workforce."

```{r}
library(tidyverse)
library(janitor)
```

# Data

```{r}
all_freq <- read_csv("data_to_share/var_frequencies_all_respondents.csv")
institutions <- read_csv("data_to_share/institutions_all.csv")
all_edu_work <- read_csv("data_to_share/education_work_all_respondents.csv")
rcd_edu_work <- read_csv("data_to_share/education_work_rcd_respondents.csv")
gender <- read_csv("data_to_share/gender_counts.csv")
ethnicity <- read_csv("data_to_share/ethnicity_counts.csv")
jobs <- read_csv("data_to_share/job_titles.csv")
facings <- read_csv("data_to_share/facings.csv")
satisfaction <- read_csv("data_to_share/satisfaction_rcd_respondents.csv")
inclusion <- read_csv("data_to_share/inclusion.csv")
```

# 3: SURVEYING THE WORKFORCE

Institution Counts

```{r}
length(unique(institutions$institution_code))
institutions %>%
  distinct(institution_code, .keep_all=TRUE) %>%
  tabyl(institution_type)
institutions %>%
  filter(!is.na(institution_code)) %>%
  count(institution_code) %>%
  summarize(median(n),
            max(n))
```

R1/R2

```{r}
institutions %>%
  filter(academic_institution) %>%
  summarize(R1_respondents = sum(institution_type == "R1")/n(),
            R1_inst = n_distinct(institution_code[institution_type == "R1"])/n_distinct(institution_code))
```


Respondents

```{r}
filter(all_freq, variable == "rcd_employed")

sum(rcd_edu_work$academic_institution, na.rm=TRUE)/sum(!is.na(rcd_edu_work$academic_institution))
```


# 4.1 Demographics

## Age

```{r}
filter(all_freq, variable=="age")
```


## Race/Ethnicity

```{r}
filter(all_freq, variable=="ethnicity1")
filter(all_freq, variable=="hispanic")
```


## Gender

```{r}
filter(all_freq, variable=="gender")
```


## US Citizenship/Residency

```{r}
filter(all_freq, variable=="us_perm")
```

## Sexual Orientation

```{r}
filter(all_freq, variable=="sex_orientation")
```

## Disability Status

```{r}
filter(all_freq, variable=="disability")
```



# 4.2 Education and Experience

## Education

```{r}
filter(all_freq, variable=="education") 
```

```{r}
filter(all_freq, variable=="education_field_count") 
```

Respondents without a PhD

```{r}
all_edu_work %>%
  filter(education != "PhD or Doctorate") %>%
  select(starts_with("education_field") & -education_field_count) %>%
  summarize(across(everything(), ~sum(.x)/n())) %>%
  pivot_longer(everything()) %>%
  arrange(desc(value))
```

Respondents with a PhD

```{r}
all_edu_work %>%
  filter(education == "PhD or Doctorate") %>%
  select(starts_with("education_field") & -education_field_count) %>%
  summarize(across(everything(), ~sum(.x)/n())) %>%
  pivot_longer(everything()) %>%
  arrange(desc(value))
```

```{r}
all_edu_work %>%
  mutate(education_related = education_related %in% c("Extremely related", "Very related")) %>%
  tabyl(education_related)
```


```{r}
all_edu_work %>%
  select(starts_with("education_field"), education_related, -education_field_count) %>%
  pivot_longer(-education_related) %>%
  mutate(education_related = education_related %in% c("Extremely related", "Very related")) %>%
  filter(value) %>%
  group_by(name) %>%
  summarize(prop_related = sum(education_related)/sum(!is.na(education_related))) %>%
  arrange(desc(prop_related))
```


## Work Experience

```{r}
median(all_edu_work$experience_work, na.rm=TRUE)
median(all_edu_work$experience_rcd, na.rm=TRUE)
```

```{r}
tabyl(all_edu_work, experience_rcd_only)
```

```{r}
median(all_edu_work$experience_rcd/all_edu_work$experience_work, na.rm=TRUE)
```

People at academic institutions with non-academic RCD work experience

```{r}
all_edu_work %>%
  filter(academic_institution) %>% 
  tabyl(experience_non_academic_rcd)
```

```{r}
tabyl(all_edu_work, experience_corporate)
```

# 4.3 Current Positions

## Employer Types and Units

```{r}
tabyl(rcd_edu_work, institution_combo_group)
tabyl(rcd_edu_work, position_group_rc)
```


## Tenure and Promotion

```{r}
median(rcd_edu_work$position_group_tenure, na.rm=TRUE)
median(rcd_edu_work$position_tenure, na.rm=TRUE)
tabyl(rcd_edu_work, position_group_tenure_cat)
```

RCD experience exceeds current group tenure?

```{r}
rcd_edu_work %>%
  mutate(other_group_rcd = experience_rcd > position_group_tenure) %>%
  tabyl(other_group_rcd)
```

Promoted

```{r}
rcd_edu_work %>%
  tabyl(promoted)
```

Only enough male and female respondents to compare (others not included below)

```{r}
gender %>%
  filter(variable=="promoted") %>%
  pivot_wider(names_from=gender, values_from=n, values_fill=0) %>%
  mutate(across(-1:-2, ~.x/sum(.x)))
  
```


```{r}
rcd_edu_work %>%
  filter(position_group_tenure_cat == "5-9") %>%
  tabyl(promoted)

rcd_edu_work %>%
  filter(position_group_tenure_cat == "10+") %>%
  tabyl(promoted)
```


## Position Types and Levels

```{r}
rcd_edu_work %>%
  tabyl(position_category)

rcd_edu_work %>%
  group_by(position_category) %>%
  summarize(rcd_experience = median(experience_rcd, na.rm=TRUE),
            supervise_staff = sum(supervise_staff, na.rm=TRUE)/sum(!is.na(supervise_staff)))
```


```{r}
gender %>%
  filter(variable=="supervise_staff") %>%
  pivot_wider(names_from=gender, values_from=n, values_fill=0) %>%
  mutate(across(-1:-2, ~.x/sum(.x)))

gender %>%
  filter(variable=="management_staff_role") %>%
  pivot_wider(names_from=gender, values_from=n, values_fill=0) %>%
  mutate(across(-1:-2, ~.x/sum(.x)))
```

```{r}
ethnicity %>%
  filter(variable=="supervise_staff") %>%
  pivot_wider(names_from=group, values_from=n, values_fill=0) %>%
  mutate(across(-1:-2, ~.x/sum(.x)))

ethnicity %>%
  filter(variable=="management_staff_role") %>%
  pivot_wider(names_from=group, values_from=n, values_fill=0) %>%
  mutate(across(-1:-2, ~.x/sum(.x)))
```

## Job Titles

```{r}
jobs %>%
  select(starts_with("job_"), -job_title_official, -job_title_alternate) %>%
  summarize(across(everything(), ~sum(.x, na.rm=TRUE)/n())) %>%
  pivot_longer(everything(), names_to="term", values_to="prop") %>%
  arrange(desc(prop))

jobs %>%
  tabyl(job_title_alternate)

jobs %>%
  summarize(sum(job_research,
                na.rm=TRUE)/n())

jobs %>%
  summarize(sum(job_compute,
                na.rm=TRUE)/n())
```

## Funding

```{r}
rcd_edu_work %>%
  filter(position_type == "Staff") %>%
  summarize(hourly = sum(!position_hourly, na.rm=TRUE)/n(),
            fulltime = sum(position_fulltime, na.rm=TRUE)/n())

rcd_edu_work %>%
  filter(position_type == "Staff") %>%
  tabyl(position_funding)
```


## Worksite

```{r}
rcd_edu_work %>%
  tabyl(onsite_remote)
```


## Responsibilities

```{r}
rcd_edu_work %>%
  tabyl(position_scope) %>%
  arrange(desc(percent))

rcd_edu_work %>%
  summarize(full_rcd = sum(position_percent_rcd == 100, na.rm=TRUE)/n(),
            rcd_76_99 = sum(position_percent_rcd > 75 & position_percent_rcd < 100, na.rm=TRUE)/n(),
            rcd_half = sum(position_percent_rcd >= 50, na.rm=TRUE)/n())

tabyl(rcd_edu_work, supervise_staff)

tabyl(rcd_edu_work, supervise_student_workers)
```



# 4.4 Position Domains

```{r}
facings %>%
  tabyl(facings_familiar)

facings %>%
  tabyl(facings_number)

facings %>%
  select(ends_with("majority")) %>%
  rowwise() %>%
  mutate(majority = sum(c_across(everything()), na.rm=TRUE)) %>%
  ungroup() %>%
  summarize(sum(majority, na.rm=TRUE)/n())

facings %>%
  select(ends_with("majority")) %>%
  summarize(across(everything(), ~sum(.x, na.rm=TRUE)/sum(!is.na(facings_researcher_majority))))

facings %>%
  select(facings_researcher:facings_strategy) %>%
  summarize(across(everything(), median, na.rm=TRUE))
```


# 4.5 Position and RCD Field Satisfaction

## Work Valued

```{r}
satisfaction %>%
  mutate(valued = valued_peers %in% c("Always", "Most of the time"),
         valued = ifelse(is.na(valued_peers), NA, valued)) %>%
  tabyl(valued)

satisfaction %>%
  mutate(valued = valued_manager %in% c("Always", "Most of the time"),
         valued = ifelse(is.na(valued_manager), NA, valued)) %>%
  tabyl(valued)

satisfaction %>%
  mutate(valued = valued_researchers %in% c("Always", "Most of the time"),
         valued = ifelse(is.na(valued_researchers), NA, valued)) %>%
  tabyl(valued)

satisfaction %>%
  mutate(valued = valued_org %in% c("Always", "Most of the time"),
         valued = ifelse(is.na(valued_org), NA, valued)) %>%
  tabyl(valued)
```


```{r}
satisfied_valued <- function(variable) {
  satisfaction %>%
    mutate(valued = {{variable}} %in% c("Always", "Most of the time"),
           valued = ifelse(is.na({{variable}}), NA, ifelse(valued, "Valued", "Not Valued")),
           satisfied = rcd_satisfaction %in% c("Extremely satisfied", "Somewhat satisfied"),
           satisfied = ifelse(is.na(rcd_satisfaction), NA, ifelse(satisfied, "Satisfied", "Not Satisfied"))) %>%
    tabyl(valued, satisfied) %>%
    adorn_percentages()
}
satisfied_valued(valued_peers)
satisfied_valued(valued_manager)
satisfied_valued(valued_researchers)
satisfied_valued(valued_org)
```

## Opportunities

```{r}
satisfaction %>%
    mutate(satisfied = rcd_satisfaction %in% c("Extremely satisfied", "Somewhat satisfied"),
           satisfied = ifelse(is.na(rcd_satisfaction), NA, ifelse(satisfied, "Satisfied", "Not Satisfied"))) %>%
    tabyl(satisfied)

satisfaction %>%
    mutate(opp = rcd_opportunities %in% c("Strongly agree", "Somewhat agree"),
           opp = ifelse(is.na(rcd_opportunities), NA, ifelse(opp, "Strongly/Somewhat Agree", "Not agree"))) %>%
    tabyl(opp)

satisfaction %>%
    mutate(opp = rcd_opportunities %in% c("Strongly agree", "Somewhat agree"),
           opp = ifelse(is.na(rcd_opportunities), NA, ifelse(opp, "Strongly/Somewhat Agree", "Not agree"))) %>%
    tabyl(promoted, opp) %>%
    adorn_percentages()

satisfaction %>%
  tabyl(rcd_future)

mean(satisfaction$rcd_recommend, na.rm=TRUE)

sum(satisfaction$rcd_recommend >= 8, na.rm=TRUE)/sum(!is.na(satisfaction$rcd_recommend))
```


```{r}
satisfaction %>%
  mutate(no_future = !rcd_opportunities %in% c("Strongly agree", "Somewhat agree") | 
           !rcd_future %in% c("Strongly agree", "Somewhat agree"),
         applied_6months = any_last_applied %in% c("Within the last month","1-6 months ago")) %>%
  tabyl(no_future, applied_6months) %>%
  adorn_percentages()
```



## Included

```{r}
inclusion
```


